Reasoning about a Capability Machine with Local Capabilities\\ Provably Safe Stack and Return Pointer Management (without OS Support).

### 2017-02-08 csf-2017/paper.tex
text about VR, readCond, and writeCond

\begin{figure}[htbp]
  \centering
  \begin{align*}
  & \writeCond{}(\iota,\gl)(W) =  \\
  & \left\{ \npair{(\start,\addrend)} \; \middle| \;
    \begin{aligned}
      & \exists r \in \var{localityReg}(g,W) \ldotp \\
      & \;\; \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp \\
      & \;\;\;\; W(r)\nsupsim[n-1] \iota_{\start',\addrend'} \text{\footnotesize{ and}}\\
      & \;\;\;\; W(r) \text{\footnotesize{ is address-stratified}} 
    \end{aligned} \right\}\\
  & \readCond{}(\gl)(W) =  \\
  & \left\{ \npair{(\start,\addrend)} \; \middle| \;
    \begin{aligned}
      & \exists r \in \var{localityReg}(g,W) \ldotp \\
      & \;\; \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp \\
      & \;\;\;\; W(r)\nsubsim[n] \iota_{\start',\addrend'}^\pwl 
    \end{aligned} \right\}\\
  & \execCond{}(\gl)(W) = \\
  & \left\{ \npair{(\var{perms},\start,\addrend)} \; \middle| \;
    \begin{aligned}
      & \forall n' < n, W' \future W \ldotp \\
      & \;\; \forall a \in [\start,\addrend], \perm \in \var{perms} \ldotp \\
      & \;\;\;\; \npair[n']{((\perm,\gl),\start,\addrend,\addr)} \in \stder(W')\\
    \end{aligned} \right\} \\
  & \entryCond{}(\gl)(W) = \\
  & \left\{ \npair{(\start,\addrend,\addr)} \; \middle| \;
    \begin{aligned}
 &  \forall n' < n \ldotp \forall W' \future W \ldotp\\
      & \quad \npair[n']{((\exec,\gl),\start,\addrend,\addr)} \in \stder(W')
    \end{aligned} \right\} \\
  & \quad \text{where } \gl = \local \Rightarrow \future = \futurewk \\
  & \quad \text{and } \gl = \glob \Rightarrow \future = \futurestr
\end{align*}
  \begin{align*}
    &\iota = (v,s,\phi_\pub,\phi,H) \text{\footnotesize{ is address-stratified iff}}\\
    &\qquad\begin{multlined}
      \forall s', \hat{W}, n, \ms, \ms' \ldotp \\
      \npair{\ms},\npair{\ms'} \in H~ s'~ \hat{W} \Rightarrow \\
      \dom(\ms) = \dom(\ms') \wedge \\
      \forall \addr \in
      \dom(\ms)\ldotp \npair{\ms\update{\addr}{\ms'(\addr)}} \in H~ s'~ \hat{W}
    \end{multlined}
  \end{align*}

@@

\subsubsection{Observation relation}
\todo[inline]{write something about single orthogonal closure?}

@@

\subsubsection{Expression and register-file relation}
The expression relation, $\stder$, is a world-indexed predicate that
contains all the words that safely can be put in the $\pcreg$-register
of a register-file that only gives access to capabilities that are capability
safe.

@@

\subsubsection{Value relation}
The value relation, $\stdvr$, contains all the words that are safe to
give away to an adversary. The conditions a capability must satisfy to
be safe to give to an adversary depends on what permissions it
gives. We there for define four conditions

Now define the value relation as follows:

>>>>

\begin{figure}[htbp]
  \centering
  \begin{align*}
  & \readCond{}(\gl)(W) =  \\
  & \left\{ \npair{(\start,\addrend)} \; \middle| \;
    \begin{aligned}
      & \exists r \in \var{localityReg}(g,W) \ldotp \\
      & \;\; \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp \\
      & \;\;\;\; W(r)\nsubsim[n] \iota_{\start',\addrend'}^\pwl 
    \end{aligned} \right\}\\
  & \writeCond{}(\iota,\gl)(W) =  \\
  & \left\{ \npair{(\start,\addrend)} \; \middle| \;
    \begin{aligned}
      & \exists r \in \var{localityReg}(g,W) \ldotp \\
      & \;\; \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp \\
      & \;\;\;\; W(r)\nsupsim[n-1] \iota_{\start',\addrend'} \text{\footnotesize{ and }}\\
      & \;\;\;\; W(r) \text{\footnotesize{ is address-stratified}} 
    \end{aligned} \right\}\\
  & \execCond{}(\gl)(W) = \\
  & \left\{ \npair{(\var{perms},\start,\addrend)} \; \middle| \;
    \begin{aligned}
      & \forall n' < n, W' \future W \ldotp \\
      & \;\; \forall a \in [\start,\addrend], \perm \in \var{perms} \ldotp \\
      & \;\;\;\; \npair[n']{((\perm,\gl),\start,\addrend,\addr)} \in \stder(W')\\
    \end{aligned} \right\} \\
  & \entryCond{}(\gl)(W) = \\
  & \left\{ \npair{(\start,\addrend,\addr)} \; \middle| \;
    \begin{aligned}
 &  \forall n' < n \ldotp \forall W' \future W \ldotp\\
      & \quad \npair[n']{((\exec,\gl),\start,\addrend,\addr)} \in \stder(W')
    \end{aligned} \right\} \\
  & \quad \text{where } \gl = \local \Rightarrow \future = \futurewk \\
  & \quad \text{and } \gl = \glob \Rightarrow \future = \futurestr
\end{align*}
  \begin{align*}
    &\iota = (v,s,\phi_\pub,\phi,H) \text{\footnotesize{ is address-stratified iff}}\\
    &\qquad\begin{multlined}
      \forall s', \hat{W}, n, \ms, \ms' \ldotp \\
      \npair{\ms},\npair{\ms'} \in H~ s'~ \hat{W} \Rightarrow \\
      \dom(\ms) = \dom(\ms') \wedge \\
      \forall \addr \in
      \dom(\ms)\ldotp \npair{\ms\update{\addr}{\ms'(\addr)}} \in H~ s'~ \hat{W}
    \end{multlined}
  \end{align*}

@@

@@

The expression relation, $\stder$, is a world-indexed predicate that
contains all the words that safely can be put in the $\pcreg$-register
of a register-file that only gives access to capabilities that are capability
safe.

@@

The \emph{value relation}, $\stdvr$, defines which words are capability
safe. A word is capability safe if it cannot be used to break the
protocols defined by the world. Data is always capability safe because
it provides no authority that can be used to break the
invariants. Capabilities gives the authority to manipulate memory and
potentially break memory protocols, so they need to satisfy certain
conditions to be capability safe. We define a condition for each kind
of permission a capability can have. These permission-based conditions
can be found in \figurename~\ref{fig:perm-conds}.

The condition for capabilities with read permission is
$\readCond{}$. This condition makes sure the read capability can only
be used to read capability safe words. To make sure this is the case,
the region that governs the part of memory the capability grants
authority over must at least require all the words to be in the value
relation.
Generally speaking, global capabilities should only depend on
permanent regions. The function $\var{localityReg}(g,W)$ projects out
all active regions when $g$ is local, but when $g$ is global, then it
is just the permanent regions.

For capabilities with a write permission, $\writeCond{}$ must be
satisfied for the capabilities range of authority. Generally speaking,
write capabilities may be used to write any word, so the region that
governs the part of memory the write capability has authority over
must allow any capability safe word.
This is of course not entirely true because of the distinction between
write and permit-write-local capabilities. For write capabilities, the
region must require the contents of memory to be non-local and
capability safe words. For permit-write-local capabilities, the
non-local requirement is omitted.

The write condition uses $\var{localityReg}$ to pick an appropriate
region in the same way the read condition does.

The condition for the region to be address-stratified is there for
technical reasons.\lau{Dominique, you probably disagree with this, so
  please feel free to add something.}

The definition of the value relation in \figurename~\ref{fig:logrel}
is defined in terms of the permission-based conditions.
